A Data-Driven Framework for Neural Field Modeling

### 2010-12-29 ltx/first_order_framework/response_to_reviews.tex

    \section{Reviewer 1}

        We first adress the reviewer's main concerns, followed by the reviewer's additional comments.

    \subsection{Main Concerns}

    \begin{enumerate}
        \item Do not use real data for the validation of their approach
        \item Assume that a few kernel basis functions accurately represent the actual connectivity kernel, which might not be the case when using real data;
        \item Do not consider the EEG/MEG lead field when constructing their mapping between the membrane voltage and the electrophysiological data (observation function)
	\emph{We interested in intracranial EEG, and maybe we can say the framework can be altered in the case of scalp EEG by considering the lead field in the mapping between the membrane voltage and the electrophysiological data, resulting in different observation equation (if  non-linear we need unscented transform on observation and if not the state estimation remains unchanged.)}
    \end{enumerate}

### 2010-12-29 ltx/first_order_framework/response_to_reviews.tex

    \section{Reviewer 1}

        We first adress the reviewer's main concerns, followed by the reviewer's additional comments.

    \subsection{Main Concerns}

    \begin{enumerate}
        \item Do not use real data for the validation of their approach
        \item Assume that a few kernel basis functions accurately represent the actual connectivity kernel, which might not be the case when using real data;
        \item Do not consider the EEG/MEG lead field when constructing their mapping between the membrane voltage and the electrophysiological data (observation function)
	\emph{We interested in intracranial EEG, and maybe we can say the framework can be altered in the case of scalp EEG by considering the lead field in the mapping between the membrane voltage and the electrophysiological data, resulting in different observation equation (if  non-linear we need unscented transform on observation and if not the state estimation remains unchanged.) This is from PLoS review which might be useful: In much of the previous papers cited the EEG/MEG lead field is
considered. This is different of course for scalp electrodes than for intracranial electrodes but should be explicitly spelled out because this factor cannot be ignored. In fact the observation model is not
quite realistic, without lead field and without amplifier characteristics.}
    \end{enumerate}

### 2010-12-30 ltx/first_order_framework/response_to_reviews.tex

\documentclass{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{anysize} %<-for setting margins
\marginsize {1.6cm}{1.6cm}{2.0cm}{2cm}
\usepackage{color}
\newcommand{\dean}[1]{\textsf{\emph{\textbf{\textcolor{red}{#1}}}}}

\title{Response To Reviews - A Data-Driven Framework for Neural Field Modeling}
\author{D. Freestone, P. Aram, M. Dewar, K. Scerri, D. Grayden, V. Kadirkamanathan}

@@

\begin{document}
    \maketitle

    \section{Reviewer 1}

        We first address the reviewer's main concerns, followed by the reviewer's additional comments.

@@

    \subsection{Main Concerns}

    \begin{enumerate}
        \item Do not use real data for the validation of their approach

	\emph{The authors appreciate that the only way to truly validate the proposed framework is with real data. However, before validating the method on real data we believe that it is important to first validate the method artificial data, where the true and estimated parameters can be compared. Given the space limitations on the article, it is not feasible to show examples using real data in this paper. Another paper demonstrating the application of the method with data recorded from a Utah array implanted in an epilepsy surgical patient is currently in preparation. In an effort to make the method employed in the paper more explicit, we have added ``To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters'' to the introduction.}

        \item Assume that a few kernel basis functions accurately represent the actual connectivity kernel, which might not be the case when using real data;

	\emph{The examples presented in the paper use the Mexican hat connectivity kernel. The framework is also valid for other connectivity kernels provided they are homogeneous and that they can be approximated by a set of isotropic basis functions, parametrized by a coefficient vector $\mathbf{\theta}$. The mathematical formulation does not restrict the shape of the kernel, where anisotropic (we assume isotropy of the kernel basis functions to make the simplification in equation~18) arbitrary shapes can be represented. Furthermore, the kernel basis functions need not be a Gaussian shape, but can be, for example, Laplacian to describe the ``wizard hat'' connectivity or B-spline functions to describe a kernel with a more compact support. For more complex basis functions some of the analytic derivations described in the appendix may have to be performed numerically. We have added the above statements to the methods section to be more explicit in the subtleties of the method.}

        \item Do not consider the EEG/MEG lead field when constructing their mapping between the membrane voltage and the electrophysiological data (observation function)
	\emph{We interested in intracranial EEG, and maybe we can say the framework can be altered in the case of scalp EEG by considering the lead field in the mapping between the membrane voltage and the electrophysiological data, resulting in different observation equation (if  non-linear we need unscented transform on observation and if not the state estimation remains unchanged.) This is from PLoS review which might be useful: In much of the previous papers cited the EEG/MEG lead field is
considered. This is different of course for scalp electrodes than for intracranial electrodes but should be explicitly spelled out because this factor cannot be ignored. In fact the observation model is not
quite realistic, without lead field and without amplifier characteristics.}
    \end{enumerate}

@@

    \subsection{Additional Comments}

    \begin{enumerate}
        \item The discrepancy of the results in Experiment II where the weights of the basis functions are not in agreement with the actual parameters while the reconstructed kernels seem satisfying.
        \item The use of Euler discretisation and how this affects the system's behavior.
        \item The derivation of Equations (30) and (31) [spacing and width of basis functions] and make them more transparent for the general audience.
        \item The existing literature on inversion schemes for similar models of brain activity.
        \item In section 3, Table 6 is mentioned however there is no such table in the ms.
    \end{enumerate}

@@

    \section{Reviewer 2}

    We thank the author for their encouraging review. Each point raised is dealt with in turn.

\begin{enumerate}
    \item The handling of the (to borrow a term from numerical weather forecasting) `sub-grid scale physics' is inadequate for nervous systems (see, e.g., equation 8 [The nonlinear IntegroDifference Equation]). I do very much appreciate the spectral estimation of spatial and temporal scales. But the model errors are lumped into an iid model error term that is not realistic. For instance, the flow of potassium in the extracellular spaces has a known dynamic that will affect network excitability, generating important sub-grid dynamics that is anything but iid. See, for instance, the example in Ullah et al PLoS Comp Biol 2010 where without such dynamics, state tracking and parameter estimation for a neural system fails. A beautiful overview of sub-grid effects in data assimilation is found in Kalnay's 2003 book. I would very much value the authors comments on these issues. I guess were I to respond to this question, that I might consider extending the model error term to an autoregressive model. But I will let the authors answer this question.\\
    \emph{talk about model order}
\item Although we all know the 1970s version of Wilson and Cowan's field equations, there was an intriguing stochasticization of these equations by Benayoun et al PloS Comp Biol 2010. I and others will value the authors' thoughts of how a more palatable stochasticization of these field equations might influence the future development of the data-assimilation framework presented in this paper.
\item This is NeuroImage, not the Physical Review. Although most of this framework involves mathematics that is reasonable to follow for typical readers of this journal, I would suggest that the authors explain and expand a bit on the introduction of Green's functions where introduced in equation 6.
\item What is the relation of the substitution of the temporal and spatial basis functions as employed in this paper to Galerkin's projection? How do the two techniques differ? If this is close, the non-local case for a Galerkin projection (typically used in diffusive processes) has not, to my knowledge, been done. Certainly not in a way that is obvious for neural field theory. Perhaps this is my ignorance. This seems important to clarify for the rest of us.
\item In the isotropic and translationally invariant case examined here, would a Fourier basis for temporal and spatial structure have been a formally correct alternative?\\
\emph{Gaussians has semi-compact support but Fourier basis functions are global functions. Gaussian can be easily extended to higher dimension (don't know about the Fourier basis functions), any orthonormal basis functions like Fourier basis functions simplify our equations, $\Gamma$ becomes identity matrix, but using Gaussian basis functions equations are in their generic form.}
\item In equation 29, what would a reasonable $v_c$ be for a real neuronal system (mammalian) given the number of neurons and scale of electrodes? I am concerned that it might be much higher than any of us wish to consider. One would perhaps be saved by the filter characteristics of brain (e.g. Phys. Rev. E 73, 051911 (2006)).
\item The two covariance inflation parameters in equation 41 make me cringe. How are we best to adjust these fudge factors?
\emph{The positive constant $\alpha$ determines the spread of the sigma points around $\mathbf{\bar x}$ which can be made arbitrary small (e.g. $0<\alpha\le 10^{-4} $) to minimise higher order effect (cite Julier (The Scaled Unscented Transformation) and  Haykin (Kalman filtering and neural network) , and  $\kappa$ is an additional scaling parameter which is usually set to $3-n_x$. cite Julier A New Approach for Filtering Nonlinear Systems}

### 2010-12-30 ltx/first_order_framework/LaTeX_Manuscript.tex
added to reviewer comments

In this paper, we demonstrate for the first time how an intracortical connectivity kernel can be inferred from data, based on a variant of the \citet{Wilson1973} neural field model. This work provides a fundamental link between the theoretical advances in neural field modeling and electrophysiological data. The paper proceeds by first describing the continuum neural field equations that are used as the cortical model. Then a finite-dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. The next section establishes conditions, using spatial frequency analysis, for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in the following section. The results for the spatial frequency analysis and parameter estimation are then presented. Finally, the implications and limitations of this framework are discussed along with planned future developments.

>>>>

In this paper, we demonstrate for the first time how an intracortical connectivity kernel can be inferred from data, based on a variant of the \citet{Wilson1973} neural field model. This work provides a fundamental link between the theoretical advances in neural field modeling and electrophysiological data. \dean{Dean (for reviewer 1 comment 1) (this is copied from the abstract): To illustrate the estimation framework, data is generated using the neural field equations incorporating modeled sensors enabling a comparison between the estimated and true parameters.} The paper proceeds by first describing the continuum neural field equations that are used as the cortical model. Then a finite-dimensional neural field model is derived. The model is reduced by approximating the neural field using a set of continuous basis functions, weighted by a finite dimensional state vector. The next section establishes conditions, using spatial frequency analysis, for both sensor and basis function spacing and width, such that the dominant dynamics of the neural field can be represented by the reduced model. The state and parameter estimation procedure is described in the following section. The results for the spatial frequency analysis and parameter estimation are then presented. Finally, the implications and limitations of this framework are discussed along with planned future developments.
